Applying gamification to developer practices

### 2013-09-07 paper_id1.tex
not sure what changed today

\section{Study Design}

@@

To setup the game for a practice, the Blaze tool provides an XML configuration file where the researcher can configure Blaze to categorize and assign points to the commands that are part of a software enginering practice.  Blaze allows the researcher to define multiple command category levels.  Thus we can categorize commands that are navigation then further classify them into structured and unstructured navigation.  We categorize commands as structured navigation when they allow the developer to locate a code element following the program structure from another element.  For example, Navigate To \textit{Ctrl+,}, Go To Definition \textit{F12}, and Find All References \textit{Ctrl+K,R} are commands that follow structure of the code, which we classify as structured navigation.
Commands that we considered as unstructured navigation include scrolling through a file using keys or the scroll-bar, opening files from the Solution Explorer, Find commands (e.g. Find in Files), Go-to line number, and children of these commands (e.g. Find Next).

@@

\section{Results}
The following hypotheses interest us in the current study.  
1. Developers using blaze use more structured navigtion when they receive instant feedback about their practices
2. Developers find the competitive aspect of blaze encourages them to pay more attention to navigation practices

To evaluate whether developers use more structured navigation when they receive instant feedback, we monitor the change in navigation ratio over the pilot study period.   Expecting that as developers learn about navigation commands and tools that count towards the score in the second week they will begin using the practices more.  Also expecting that noticing the leaderboard competition will encourage some developers to use the navigation practices more.  What we observed is shown in Figure \ref{fig:navigationaverage} that the structured navigation practice did not increase significantly from the beginning to the end of the study or between weeks when the additional information was revealed to them.

@@

>>>>

\section{Study and Game Design}

@@

To setup the game for a practice, the Blaze tool provides an XML configuration file where the researcher can configure Blaze to categorize and assign points to the commands that are part of a software engineering practice.  Blaze allows the researcher to define multiple command category levels.  Thus we can categorize commands that are navigation then further classify them into structured and unstructured navigation.  We categorize commands as structured navigation when they allow the developer to locate a code element following the program structure from another element.  For example, Navigate To \textit{Ctrl+,}, Go To Definition \textit{F12}, and Find All References \textit{Ctrl+K,R} are commands that follow structure of the code, which we classify as structured navigation.
Commands that we considered as unstructured navigation include scrolling through a file using keys or the scroll-bar, opening files from the Solution Explorer, Find commands (e.g. Find in Files), Go-to line number, and children of these commands (e.g. Find Next).

@@

\section{Results}

Although we translated navigation commands to points for developers, we defined the metric Navigation ratio based on Jones and Jones \cite[] as the number of structured navigation events in a day over the number of unstructured navigation events.  This metric provides a continuous variable allowing us to use statistical data of continuous data to check our hypotheses.  To create the evaluation for each wave, we took the navigation ratio for each developer in each wave.

The following hypotheses interest us in the current study.  
1. Developers using blaze use more structured navigation when they receive instant feedback about their practices
2. Developers find the competitive aspect of blaze encourages them to pay more attention to navigation practices

To evaluate whether developers use more structured navigation when they receive instant feedback, we monitor the change in navigation ratio over the pilot study period.   Expecting that as developers learn about navigation commands and tools that count towards the score in the second week they will begin using the practices more.  Also expecting that noticing the leaderboard competition will encourage some developers to use the navigation practices more.  What we observed is shown in Figure \ref{fig:navigationaverage} that the structured navigation practice did not increase significantly from the beginning to the end of the study.  Overall, we did not see a significant effect towards increasing navigation ratio for the interventions in wave 2 or wave 3.

@@

Is this a spectacular fail from a gamification effectiveness perspective?   How could this be when pre-study results indicated very positive attitudes towards the approach?  Our post pilot survey attempts to shed some light on the reasons.  One question in the post survey is how knowledgeable the developer was about the code they worked in with 3 possibilities,  they were authors of the code, they regularly maintained the code, they were unfamiliar with the code.  In this team, x\% of developers cited that they ...the code.  Another factor to consider is experience level which we also asked in the post-pilot survey.  More experienced developers may be more convinced of their effective habits and perhaps react almost with muscle memory to use structured or unstructured navigation to locate code of interest.  Even developers whose navigation ratio  was lower than average, used structured navigation occasionally as much as their more frequently using colleagues.  The variation for each individual can be seen in Figure \ref{} .
